---
title: "Food Consumption & Carbon Footprint"
output: html_notebook
---


##### Loading Libraries and data
```{r}
library(tidyverse)
library(knitr)
library(readr)
library(skimr)
library(echarts4r)
```


```{r}
#Load data
food_consumption <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-18/food_consumption.csv')
```


##### Data Overview and Summary
```{r}
#Overview of data
# Consumption is measured in kg/person/year
# Co2 Emission is measured in kg CO2/person/year
food_consumption %>%
  head(10) %>%
  kable()

```


```{r}
# Data Summary
# There are 130 unique countries and 11 food categories
skim(food_consumption)
```



```{r}
# Transforming the data
combined_consumption <- food_consumption %>% 
    pivot_longer(3:4, names_to = "feature", values_to = "value")

combined_consumption %>% 
  head() %>% 
  kable()
```



##### Distribution of Food Consumption and CO2 Emissions
```{r}
# Overall distribution of co2 and food consumption
combined_consumption %>%
  ggplot(aes(x=feature, y=value, color = feature)) +
  geom_boxplot() +
  facet_wrap(~feature, scales = "fixed") + 
  scale_y_log10() +
  labs(title = "Distribution of CO2 Emission and Food Consumption")
  
```


##### CO2 Emissions by Food Categories
```{r}
# Food categories with the highest CO2 Emissions
food_consumption %>%
  mutate(co2_eachfood = co2_emmission/ consumption) %>%
  group_by(food_category) %>% 
  summarize(avg_CO2 = mean(co2_eachfood, na.rm = T)) %>% 
  ggplot(aes(fct_reorder(food_category, avg_CO2), avg_CO2, fill = food_category)) +
  geom_col(show.legend = F) +
  coord_flip() +
  labs(x = "Food Categories", y = "Avg CO2 Emissions", title="Average CO2 Emissions by Food Category")

```
```{r}
# Lamb & Goat, Beef have significantly higher CO2 emissions compared to the other food categories
```


##### Distribution of CO2 Emissions and Consumption across all Food Categories
```{r}
# Emissions vs Consumption for all food categories
combined_consumption %>%
   ggplot() +
   geom_bar(aes(x = feature, y = value, fill = feature), stat = "identity") +
   facet_wrap(~food_category, scales = "free") +
   theme(legend.position = "bottom",
         axis.text.x = element_blank()) + 
   labs(title = "CO2 Emissions vs Consumption across Food Categories",
        x = "Features",
        fill = "Feature")
```


##### Top 5 Countries with Highest Consumption in All Categories
```{r}
# Top 5 countries in each food category with highest consumption
food_consumption %>% 
  group_by(food_category) %>% 
  top_n(consumption, n = 5) %>% 
  mutate(country = fct_reorder(country, consumption)) %>% 
  ggplot(aes(consumption, country)) +
  geom_col(aes(fill = food_category), show.legend = F) +
  facet_wrap(~food_category, scales = "free") +
  theme(axis.text.x = element_blank()) +
  labs(title = "Countries with Highest Consumption in All Food Categories",
         x = "food consumption",
         y = "")

```


##### Top 5 Countries with Highest CO2 Emissions in All Food Categories
```{r}
# Top 5 countries with highest CO2 emissions in each food category
food_consumption %>% 
  group_by(food_category) %>% 
  top_n(co2_emmission, n = 5) %>% 
  mutate(country = fct_reorder(country, consumption)) %>% 
  ggplot(aes(co2_emmission, country)) +
  geom_col(aes(fill = food_category), show.legend = F) +
  facet_wrap(~food_category, scales = "free") +
  theme(axis.text.x = element_blank()) +
  labs(title = "Countries with Highest CO2 Emissions in All Food Categories",
         x = "CO2 Emission",
         y = "")
```


##### Summary of Total Consumption and Total C02 Emissions
```{r}
consumption_map <- food_consumption %>%
  group_by(country) %>%
  summarise(total_consumption = sum(consumption),
            total_CO2emission = sum(co2_emmission)) %>%
  arrange(-total_CO2emission) %>%
  mutate(country = recode_factor(country,
                                  `USA` = "United States",
                                  `Czech Republic`= "Czech Rep.",
                                  `South Korea`= "Korea"))

consumption_map %>%
  head() %>%
  kable()

```


```{r}
consumption_map %>%
  e_charts(country) %>%
  e_map(total_CO2emission) %>%
  e_visual_map(min = 0, 
               max = 2200) %>%
  e_title("Total CO2 Emissions by Country", 
          left = "center") %>%
  e_theme("azul")
  
```
